---
title: "Lesson 4 - Ordinal Response Models"
author: "Jacob S. Lewis"
date: "2022-10-06"
output: word_document
editor_options: 
  chunk_output_type: console
---

# Libraries
```{r LIbraries}
library(haven) # Brings in data from other formats (e.g., SPSS)
library(tidyverse) # Brings in all the tidyverse packages!
library(ordinal) # Allows for pooled and multilevel ordinal response models
library(marginaleffects) # Easy plotting and generation of predicted probabilities
```

# Ordinal data
1. What are they?
2. How and why are they diff. from continuous data?
3. What does it mean to make predictions with them?
4. Predictions - bootstrapping?


# Data
What type of data use ordinal?
1. Survey responses, often (discuss "feeling thermometers")
2. Categorized data to rank things



Let's bring in the Afrobarometer data and examine evaluations of identity
1. Identity variable - Q85B
2. Urban / Rural - URBRUR


```{r Data - Afrobarometer}
# Import the Afrobarometer data from SPSS format using the `haven` package
a7 <- read_sav("Data sets/r7_merged_data_34ctry.release.sav") %>% # Bring in the data
  select(RESPNO, COUNTRY, URBRUR, Q85B) %>% # Include only a few variables
  filter(Q85B == 1 | Q85B == 2 | Q85B == 3 | Q85B == 4 | Q85B == 5) %>% # Only include values of identity b/t 1:5
  filter(!URBRUR == 460) %>% # Get rid of NA values for urban/rural
  mutate(identity = factor(Q85B,
                           levels = c(1, 2, 3, 4, 5),
                           labels = c("Only ethnic",
                                      "Mostly ethnic",
                                      "Equally ethnic/national",
                                      "Mostly national",
                                      "Only national")),
         location = factor(URBRUR,
                           levels = c(1, 2, 3),
                           labels = c("Urban", "Rural", "Semi-Urban"))) %>%
  glimpse()


```

Let's evaluate the data (Visualization)
```{r Visualize the data}

# Urban/Rural and identity
ggplot(data = a7,
       aes(x = identity,
           fill = location)) +
  geom_bar(position = "fill",
           color = "black") +
  labs(title = "Breakdown of identity and respondent location",
       x = "",
       y = "Count of individuals per localtion",
       fill = "Location") +
  theme(panel.border = element_rect(fill = NA, color = "black"))


```

What happens when we run an LPM rather than a CLM?
```{r LPM}
reg3 <- lm(as.numeric(identity) ~ location, data = a7)
summary(reg3)

marginaleffects::predictions(reg3,
                             newdata = datagrid(location = c("Rural", "Semi-Urban", "Urban"))) %>%
  ggplot(aes(x = location,
             y = predicted,
             label = sprintf("%0.2f", round(predicted, digits = 2)))) +
  geom_bar(stat = "identity",
           color = "black",
           fill = "gray") +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high)) +
  geom_text(vjust = -1.5) +
  coord_cartesian(ylim = c(1,5))
```

Alright, now let's do it properly with a CLM!
```{r CLM}

# First, we'll run a factor IV, which is equivalent to a fixed effect model
reg1 <- ordinal::clm(identity ~ location, data = a7)
summary(reg1)

## Now we can plot this out
## Approach 1 (I prefer this)
marginaleffects::predictions(reg1,
                             newdata = datagrid(location = c("Rural", "Semi-Urban", "Urban"))) %>%
  ggplot(aes(x = group,
             y = predicted,
             fill = location,
             label = sprintf("%0.2f", round(predicted, digits = 2)))) +
  geom_bar(stat = "identity",
           position = "dodge",
           color = "black") +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                position = position_dodge(0.9)) +
  geom_text(position = position_dodge(0.9),
            vjust = -1.5)

## Approach 2 - useful, but harder to read
marginaleffects::predictions(reg1,
                             newdata = datagrid(location = c("Rural", "Semi-Urban", "Urban"))) %>%
  ggplot(aes(x = group,
             y = predicted)) +
  geom_bar(stat = "identity",
           color = "black",
           fill = "gray") +
  facet_wrap(~location,
             ncol = 5) +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                position = position_dodge(0.9)) +
  theme(panel.border = element_rect(fill = NA, color = "black"),
        strip.background = element_rect(fill = NA),
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1),
        strip.text = element_text(size = 12, face = "bold"))


# Now, we'll run it as if location is continuous
reg2 <- ordinal::clm(identity ~ as.numeric(location), data = a7)
summary(reg2)

## Now we can plot this out
## Approach 1 (I prefer this)
marginaleffects::predictions(reg2,
                             newdata = datagrid(location = c(1, 2, 3))) %>%
  ggplot(aes(x = group,
             y = predicted,
             fill = as.factor(location),
             label = sprintf("%0.2f", round(predicted, digits = 2)))) +
  geom_bar(stat = "identity",
           position = "dodge",
           color = "black") +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                position = position_dodge(0.9)) +
  geom_text(position = position_dodge(0.9),
            vjust = -2)



```


