---
title: "Lesson 1 - The Basics"
author: "Jacob S. Lewis"
date: '2022-07-22'
output: html_document
editor_options: 
  chunk_output_type: console
---


# R Basics

R is a computer language designed for maths and statistics.

It has some big advantages:
1. It's super powerful.
2. It's open source and easily extensible
3. It allows you to work with all sorts of data at once
4. It's free

I'm not going to go through the whole rigamarole of introducing you to every basic R function, but we will go through some important ones.

First, R can do basic math.
```{r R Basic Math}

# We can see whether we live in Oceania or not.
2+2

# Phew! We're still living in the free world ... for now.
```

Second, R is based on the idea of **objects**. It's not like Stata or Excel, which essentially are user interfaces that wrap themselves around a file. Instead, imagine your R workspace like a desk. You can pile lots of books, papers, maps, drawings, and even 3D prints on it and then interact with all of them!

If we think about objects, we can begin looking at some core R functions.
```{r R Objects}

# Let's start by creating an object called "object" and telling R that it simply contains the letter "a". Let's start by using the "<-" function.

object <- "a"

# Now, let's confirm that the object really is just the letter "a".
object

# Now we can create another object using the "=" function. It does the same thing!
object2 = "b"

# Add them together?
object + object2

# Nope! Doesn't work

# So, how to we bring them together? Two easy ways:
# concatenate them with c()
new.object <- c(object, object2)

# Paste them together with paste()
new.object2 <- paste(object, object2, sep = ".")

# Let's use numbers
object3 = 48
object4 = 26

# We can engage in simple arithmetic with numeric objects
object3 + object4
object5 = object3 + object4
object6 = object3 / object4

# Let's do something crazy, mixing numbers and characters!
crazy <- paste(object, object3*object4, object2, sep = ":")
crazy
```

**Note:** Whenever we create objects or load packages, we are loading them into our `environment`. 

Of course, we will often have to work with large, complex objects with lots of values in them. R does this really well. It's up to us to figure out some basic commands to understand how to navigate these objects.

```{r R Complex objects}

# A simple set of objects might include a string of numbers or characters

# Let's start with numbers. We'll manually create a list (vector) by concatenating several numbers:
numbs1 <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)

# We can also automatically fill out a list of 1 to 9 using the ":" function:
numbs2 <- 1:9

# We can begin to select individual items from within the list by focusing on the position of that item. Generally, brackets "[]" are used to "open up" an object and specify something about it. If we wanted to specify that we want the third item in the list of `numbs1`, we would use:
numbs1[3]
numbs1[3:6]

# We can even do complex operations using this notation!
numbs1[4] + numbs1[8]


# We can generate some random numbers. As with many things in R, this can be done in many ways.

# Let's start by using the RANDOM UNIFORM NUMBERS function (runif()). Let's start by simply examining what this function does:
?runif

# Let's create 100 random numbers between 1 and 100. We can then run a quick histogram using R's built-in package to check whether they are uniformly distributed.
numbs3 <- runif(100, min = 1, max = 100)
# WHAT SHOULD THIS LOOK LIKE?
hist(numbs3)

# Now we can create random numbers from a NORMAL DISTRIBUTION using rnorm(). After that, we'll once again histogram it.

# Again, let's look at the function
?rnorm

numbs4 <- rnorm(100, mean = 50, sd = 10)
# WHAT SHOULD THIS LOOK LIKE?
hist(numbs4)
numbs4[36:40]
hist(numbs4[4:26])



# The more numbers we put into each of these, the more they will resemble either uniform or normal distributions.
numbs5 <- runif(10000, min =  1, max = 100)
hist(numbs5)


numbs6 <- rnorm(10000, mean = 50, sd = 10)
hist(numbs6)

```

Our data are not always numerical, however! Sometimes, we have data comprised of lists of words.
```{r R Strings}
# Let's create an easy concatenate of strings
str1 <- c("Washington", "State", "University", "Go", "Cougs!")

# The fourth word in str1 is ...
str1[4]

```

We will return to dealing with strings in a future class, but for now, this is all you need to know.

Of course, as social scientists, we generally want to look at datasets that have multiple dimensions! So, let's start by simply creating a fake dataset of students.

```{r R Creating a fake dataset and messing around with it}

# We create a dataset easily using a base function of data.frame

df <- data.frame(names = c("Roger", "Ayal", "June", "Bjorn", "Todd", 
                           "Peter", "Thabo", "Florence", "Mishka", "Ife", 
                           "Mohamed", "Juanita", "Maria", "Max", "Yael", 
                           "Dana", "Preeti", "Sarah", "Anastasia", "Dawn"),
                 grade = rnorm(20, mean = 85, sd = 10),
                 study = rnorm(20, mean = 60, sd = 15))

plot(df$study, df$grade)

grade = rnorm(20, mean = 85, sd = 10)
study = rnorm(20, mean = 60, sd = 15)

plot(study, grade)



```

# Getting some real data

While it is really useful to know how to create data frames from scratch, usually you will want to use data that actually represent something! So let's take a look at some actual data frames that we might be interested in.

**Importing data!**
R has a built in function to import data: `read.csv`
But we can also use the tidyverse solution: `readr::read_csv`
Finally, R Studio has an import dataset function in its *Environment* tab

We'll start with the base R functions. They are kind of clunky, but worth knowing.

Let's start with the ANES data: https://electionstudies.org/. You have to register. I downloaded the 2020 data. There is also a package called `anesr` which allows you to simply bring all the data right in: https://github.com/jamesmartherus/anesr

```{r Data - ANES - Doing things the OLD way}
# Let's import data THE OLD WAY
anes <- read.csv(file = "Data sets/anes/anes_timeseries_2020_csv_20220210.csv")

# The first thing we ALWAYS DO is to examine our data!
nrow(anes)
ncol(anes)
colnames(anes)
View(anes)

# We pretty obviously need to look at the codebook. Let's try to find information for the following variables: gender and intention to vote

# Gender = V201600 (1 = male, 2 = female, )
# What happens if we simply summarize?
summary(anes$V201600)

# Why does it show up that way?!
mode(anes$V201600)

# One way to look at the data is to transform the data from numeric to factorial!
summary(as.factor(anes$V201600))

# Or, we could tabulate the data using table()
table(anes$V201600)

# What does this tell us?


# Who to vote for? = V201033 (1 = Biden, 2 = Trump, 3 = Jorgensen, 4 = Hawkins)
table(anes$V201033)

# Should we look at the interaction of gender and language?
## Method 1: table
table(anes$V201600, anes$V201033) # simply tabulates data by calling the dataset and specific column (using $)

with(anes, table(V201600, V201033)) # We can specify that we are using ANES data by stating "with" first

# Or, we can attach the data to tell R that any object is actually a component of the dataset!
attach(anes)
table(V201600, V201033)
detach(anes)




# We can see that: (a) the variable names are a pain in the butt, and (b) they are often packed with values that we can't use. Let's simplify our lives.

# Create new variables

## Gender
anes$gender <- NA
anes$gender[anes$V201600 == 1] <- "Male"
anes$gender[anes$V201600 == 2] <- "Female"
anes$gender <- factor(anes$gender)
summary(anes$gender)

## Vote
anes$vote <- anes$V201033
anes$vote <- factor(anes$vote, 
                    levels = c(-9, -8, -1, 1, 2, 3, 4, 5, 11, 12),
                    labels = c("Refused", "Don't know", "Inapplicable",
                               "Joe Biden", "Donald Trump", "Jo Jorgensen",
                               "Howie Hawkins", "Other", "Don't know", "Refused"))

table(anes$gender, anes$vote)



## We can actually look at the two candidates that matter
anes$bt <- NA
anes$bt[anes$V201033 == 1] <- "Biden"
anes$bt[anes$V201033 == 2] <- "Trump"
table(anes$bt)
mode(anes$bt)

with(anes, table(gender, bt))

# Now we can check out the same thing, but using xtabs
xtabs(~gender+bt, data = anes)

anes$trump <- NA
anes$trump[anes$V201033 == 1] <- 0
anes$trump[anes$V201033 == 2] <- 1
table(anes$trump)
mode(anes$trump)

t.test(trump ~ gender, 
       alternative = "less",
       conf.level = 0.99,
       var.equal = TRUE,
       data = anes)


```

```{r Shape down ANES using tidyverse}
# Now, let's try managing our data using `tidyverse`
## Bring in the library
library(tidyverse)

# Now we will use a series of functions to 
anes2 <- readr::read_csv("Data sets/anes/anes_timeseries_2020_csv_20220210.csv") %>%
  mutate(gender = factor(V201600,
                         levels = c(-9, 1, 2),
                         labels = c("Refused", "Male", "Female")),
         vote = factor(anes$V201033, 
                    levels = c(-9, -8, -1, 1, 2, 3, 4, 5, 11, 12),
                    labels = c("Refused", "Don't know", "Inapplicable",
                               "Joe Biden", "Donald Trump", "Jo Jorgensen",
                               "Howie Hawkins", "Other", "Don't know", "Refused"))) %>%
  select(c(gender, vote)) %>%
  glimpse()


# This is an easy bar plot
ggplot(data = anes2,
       aes(x = gender, fill = vote)) +
  geom_bar()

# Let's position the bars next to each other
ggplot(data = anes2,
       aes(x = gender, fill = vote)) +
  geom_bar(position = position_dodge())

# Let's look at total proportions
ggplot(data = anes2,
       aes(x = gender, fill = vote)) +
  geom_bar(position = position_fill())

# Let's get rid of the "refused" gender, since they're so tiny
## We can do it by selecting which genders to include ....
ggplot(data = subset(anes2, gender %in% "Male" | gender %in% "Female"),
       aes(x = gender, fill = vote)) +
  geom_bar(position = position_fill())

## Or we can do it by selecting which gender NOT to include ...
ggplot(data = subset(anes2, !gender %in% "Refused"),
       aes(x = gender, fill = vote)) +
  geom_bar(position = position_fill())


# Let's mess around with the plot!
ggplot(data = subset(anes2, !gender %in% "Refused"), # Let's use the anes2 data, but only M/F
       aes(x = gender, fill = vote)) + # We'll see gender to be on the x axis and the fill color to be how they voted
  geom_bar(position = position_fill(), # Let's visualize using a bar plot and the position as "fill" to fill to 100%
           color = "black", size = 1) + # Here we make the border black and the size of the border 1
  theme(legend.position = "right", # Opening up the theme allows us to mess with the non-geom aspects of the plot
        panel.background = element_rect(fill = "gray95"), # Set the background color using element_rect
        panel.border = element_rect(color = "red", 
                                    linetype = "dashed",
                                    fill = NA, 
                                    size = 2), # Set the panel border, remember fill = NA!
        axis.text.x = element_text(size = 14, angle = 45, hjust = 1), # Mess with x axis text!
        axis.text.y = element_text(size = 12),
        axis.title.y = element_text(size = 14, face = "bold", color = "darkgray")) + # Mess with y axis text
  scale_fill_brewer(palette = "BrBG") + # Determine which color palette to use
  labs(title = "2020 Elections",
       subtitle = "Voting by gender",
       x = "",
       y = "Proportion voting (observations not equivalent",
       fill = "Candidate")

```

# Dealing with libraries / packages
```{r Introducing the here package}

## HERE ###############################
# One of the most important, but easiest, packages is here(). It makes it super easy to set your entire project at the current location of the project.
library(here)

# We can initialize here() by simply typing ...
here()

# Now all we need to do is to use the here command rather than using complex directories!
anes3 <- read.csv(file = here::here("Data sets", "anes", "anes_timeseries_2020_csv_20220210.csv"))

# This is particularly helpful when we have lots and lots of complex directories or file names.


```

Let's bring in some other data
```{r Data - VDEM - Using tidyverse}
# Let's bring in the V-Dem data (a major dataset from Gothenburg that basically replaces the Polity IV data).

# Let's look at countries, years, polyarchy (their measurement of democracy) AND let's look at some other potential interesting variables (GDP, civil war, etc.)

vdem <- read_csv(file = here::here("Data sets", "vdem", "vdem" ,"V-Dem-CY-Full+Others-v12.csv")) %>%
  select(c(country_name, year, e_area, e_regiongeo, v2x_polyarchy, e_gdp)) %>%
  filter(year >= 1900) %>%
  mutate(period = if_else(year >= 1900 & year < 1914, "Pre-WWI",
                          if_else(year >=1914 & year < 1919, "WWI",
                                  if_else(year >= 1919 & year < 1939, "Inter-War",
                                          if_else(year >= 1939 & year < 1946, "WWII",
                                                  if_else(year >= 1946 & year < 1992, "Cold War", "Current era")))))) %>%
  mutate(period = factor(period, levels = c("Pre-WWI",
                                            "WWI",
                                            "Inter-War",
                                            "WWII",
                                            "Cold War",
                                            "Current era"))) %>%
  glimpse()




# Let's check out what our variables look like.
## Look at the distribution of the polyarchy variable using a density plot
## Density plots are basically smoothed out histograms, but rather than showing counts, they show us the kernel density estimation - basically the probability 
ggplot(data = vdem,
       aes(x = v2x_polyarchy)) +
  geom_density()

# Let's break it out by period
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           color = period)) +
  geom_density()

# Let's go crazy
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           color = period,
           linetype = period)) +
  geom_density()

# Let's go ... even crazier.
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           color = period,
           linetype = period,
           size = period)) +
  geom_density()


# We can make this look a little nicer
ggplot(data = vdem,
       aes(x = v2x_polyarchy)) +
  geom_density(fill = "white",
               size = 1) +
  theme(panel.border = element_rect(fill = NA,
                                    color = "black",
                                    size = 1)) +
  labs(title = "Density of the V-Dem Polyarchy variable",
       x = "V-Dem Polyarchy variable",
       y = "Density")

# It might be that we should simply use a histogram
ggplot(data = vdem,
       aes(x = v2x_polyarchy)) +
  geom_histogram(color = "black") +
  theme(panel.border = element_rect(fill = NA,
                                    color = "black",
                                    size = 1)) +
  labs(title = "Histogram of the V-Dem Polyarchy variable",
       x = "V-Dem Polyarchy variable",
       y = "Count")

# Let's break it out by period
ggplot(data = vdem,
       aes(x = v2x_polyarchy)) +
  geom_histogram(color = "black") +
  theme(panel.border = element_rect(fill = NA,
                                    color = "black",
                                    size = 1)) +
  labs(title = "Histogram of the V-Dem Polyarchy variable",
       x = "V-Dem Polyarchy variable",
       y = "Count") +
  facet_wrap(~period)

# But there's something kind of confusing about this - what?

ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           fill = period)) +
  geom_histogram(color = "black",
                 position = "fill") +
  theme(panel.border = element_rect(fill = NA,
                                    color = "black",
                                    size = 1)) +
  labs(title = "Histogram of the V-Dem Polyarchy variable",
       x = "V-Dem Polyarchy variable",
       y = "Count")




# Now we can start to interact two variables together
## How about a scatterplot of polyarchy and GDP?
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           y = e_gdp)) +
  geom_point()

# Well, that looks insane. Let's log our GDP.
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           y = log(e_gdp + 1))) +
  geom_point()

# Okay, that looks better. But we can start to examine the data a bit more.
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           y = log(e_gdp + 1),
           color = year)) +
  geom_point()


# Let's differentiate the colors a bit
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           y = log(e_gdp + 1),
           color = year)) +
  geom_point() +
  scale_color_continuous(type = "viridis")

# What abotu simplifying?
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           y = log(e_gdp + 1),
           color = year)) +
  geom_point() +
  scale_color_binned(type = "viridis")


# Now let's break it out by time period
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           y = log(e_gdp + 1),
           color = year)) +
  geom_point() +
  scale_color_binned(type = "viridis") +
  facet_wrap(~period)

# Now spiff it up.
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           y = log(e_gdp + 1),
           color = year)) +
  geom_point() +
  scale_color_binned(type = "viridis") +
  facet_wrap(~period) +
  theme(panel.border = element_rect(fill = NA, color = "black"),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold", size = 10)) +
  labs(title = "Examining Polyarchy and GDP by era",
       x = "V-Dem Polyarchy variable",
       y = "Log of GDP",
       color = "Year")


# It might be helpful to add a regression line!
ggplot(data = vdem,
       aes(x = v2x_polyarchy,
           y = log(e_gdp + 1),
           color = year)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y~x) + # <- an easy way to add the line :)
  scale_color_binned(type = "viridis") +
  facet_wrap(~period) +
  theme(panel.border = element_rect(fill = NA, color = "black"),
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face = "bold", size = 10)) +
  labs(title = "Examining Polyarchy and GDP by era",
       x = "V-Dem Polyarchy variable",
       y = "Log of GDP",
       color = "Year")


# Someone tell me about these plots. What do we learn?
# General takeaway?
# Skedasticity?
# WWMSS? What would Mike Salamone Say?

```

Having done some core visualization, let's try a quick regression (OLS)
```{r OLS - VDEM}
# Why is it appropriate to use OLS for this?

ols.vd.1 <- lm(log(e_gdp+1) ~ v2x_polyarchy,
               data = vdem)

# Look quickly at this and someone tell me what it means .. but be careful, there's a trap (think log-lin)
summary(ols.vd.1)

# We might also wonder about how robust these standard errors are, for two reasons
# A - robustness against heteroskedasticity

# So let's load in some packages to deal with them

library("lmtest")
library("sandwich")

coeftest(ols.vd.1, 
         vcov = vcovHC,
         type = "HC0")

# And what the heck is HC0"!? There are different HC estimators. Honestly, I don't know much about them. HC0 is the OG.

# There are many more post-estimation tests to run here: https://www.r-econometrics.com/methods/hcrobusterrors/

# But what about clustering? Why should we care about clustering standard errors?




# We can cluster our standard errors around the era, for example:
coeftest(ols.vd.1, 
         vcov = vcovHC,
         type = "HC0",
         df = 5,
         cluster = ~period)



# This is also an excellent explainer! https://evalf21.classes.andrewheiss.com/example/standard-errors/

# We can also cluster our standard errors around the region!
coeftest(ols.vd.1, 
         vcov = vcovCL,
         type = "HC0",
         df = 5,
         cluster = ~e_regiongeo)


```



