---
title: "Class 5 - Count Models"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r Libraries}
# Bring in our normal libraries
library(tidyverse)
library(marginaleffects)
library(modelsummary)

# Bring in some statistical libraries
library(pscl) # The "Political Science Computational Library" library
library(MASS) # Modern Applied Statistics with S

# Bring in the HERE library

library(here)
here()
```

# Count models versus Linear Regression
What types of questions might we ask that would work for count models:
1. How many protests occur in response to a new policy change?
2. How many deaths occurred in different wars?

Why should we use a count model rather than a linear regression?

Poisson distribution

In a `Poisson distribution`, we anchor the left-hand of the distribution to 0 and adjust the `lambda`, which is the average count of things at a given time/place.
```{r Draw multiple Poisson distributions}

# Set up the data frame
p_dat <- map_df(1:10, ~ tibble(
  l = paste(.),
  x = 0:20,
  y = dpois(0:20, .)
))


# Build Normal distributions
n_dat <- map_df(1:10, ~ tibble(
  l = paste(.),
  x = seq(0, 20, by = 0.001),
  y = dnorm(seq(0, 20, by = 0.001), ., sqrt(.))
))


# Use ggplot2 to plot
ggplot(n_dat, aes(x, y, color = factor(l, levels = 1:10))) +
  geom_line() +
  geom_point(data = p_dat, aes(x, y, color = factor(l, levels = 1:10))) +
  labs(color = "Lambda:") +
  theme_minimal()
```


## Poisson models

**Assumptions**
1. The response variable is a count per unit of time/space in a Poisson distribution
2. Observations must be independent of one another (this can be a problem in social science!)
3. The mean of the variable must equal its variance (this can also be a problem!!!)
4. The log of the mean rate must be a linear function of the IV


# Using the UCDP/PRIO Global Events Data
Let's take a look at a few different potential count models we could run!

**First**, perhaps we want to know about deaths.

Import the data
```{r Bring in GED data for attacks and deaths}
# Import the data
ged <- read_csv(here::here("Data sets", "ged221-csv", "GEDEvent_v22_1.csv")) %>%
  as_tibble() %>%
  mutate(date_start = lubridate::ymd(date_start),
         date_end = lubridate::ymd(date_end),
         type_of_violence = factor(type_of_violence,
                                   levels = 1:3,
                                   labels = c("State-based violence",
                                              "Non-state conflict",
                                              "One-sided violence")),
         duration = as.numeric(date_end - date_start)+1) %>%
  glimpse()
```

First, run this using OLS (LPM)
```{r LPM}
# Run the model
lpm1 <- lm(deaths_civilians ~ duration, data = ged)


# Generate predictions
marginaleffects::predictions(lpm1,
                             newdata = datagrid(duration = c(1:20))) %>%
  ggplot(aes(x = duration,
             y = predicted)) +
  geom_line() +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high))


```

Okay, now run this using a Poisson model
```{r Poisson}
# Run a Poisson count model
p1 <- glm(deaths_civilians ~ duration, 
          family = poisson,
          data = ged)


# First, we can bootstrap the predictions
## Bootstrap "as-is"
ged$pred <- exp((coef(p1)[[1]] * 1) +
                  (coef(p1)[[2]] * ged$duration))

## From this, we can extract a mean (m) and a standard error (se)
mean(ged$pred, na.rm = TRUE)
sd(ged$pred, na.rm = TRUE)/sqrt(nrow(ged))

## Bootstrap for values 1 through 20
preds <- map_df(1:20,
                ~ tibble(
                  duration = (.),
                  dur_m = mean(exp((coef(p1)[[1]] * 1) + (coef(p1)[[2]] * duration)), na.rm = TRUE)))

# Now, we can use the marginaleffects package
marginaleffects::predictions(p1,
                             newdata = datagrid(duration = c(1:20))) %>%
  ggplot(aes(x = duration,
             y = predicted)) +
  geom_line() +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high))

```

Check these out against one another
```{r Place predictions against one another}
data.frame(mode = c(rep("LPM", 20), rep("GLM", 20)),
                    duration = rep(1:20, 2),
           predicted = c(marginaleffects::predictions(lpm1,
                             newdata = datagrid(duration = c(1:20)))[,3],
                         marginaleffects::predictions(p1,
                             newdata = datagrid(duration = c(1:20)))[,3]),
           conf.low = c(marginaleffects::predictions(lpm1,
                             newdata = datagrid(duration = c(1:20)))[,7],
                         marginaleffects::predictions(p1,
                             newdata = datagrid(duration = c(1:20)))[,7]),
           conf.high = c(marginaleffects::predictions(lpm1,
                             newdata = datagrid(duration = c(1:20)))[,8],
                         marginaleffects::predictions(p1,
                             newdata = datagrid(duration = c(1:20)))[,8])) %>%
  ggplot(aes(x = duration,
             y = predicted,
             color = mode)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high))
```


**Second**, what about number of attacks in a given country?

First, we'll need to modify the GED data
```{r Create a tally of attacks in a given country per year}
# Let's tally all of this down to the year
ged2 <- ged %>%
  arrange(country, date_start) %>%
  filter(type_of_violence  %in% "Non-state conflict") %>%
  group_by(country, year) %>%
  tally()

## Take a look at this:
ged2

## We can also create a linkage between GED country names and country_id
ged3 <- ged %>%
  dplyr::select(country, country_id) %>%
  group_by(country) %>%
  summarise(country_id = mean(country_id))

```

Alright, so what is the problem with this?
WE ONLY RECORD OBSERVATIONS WHERE COUNTRIES AND YEARS HAVE CONFLICT

How do we deal with this?

```{r Create blank index and fill it in}
# First, extract each individual country from the GED data (this still has a problem, but okay)
countries <- unique(ged$country)

# Now we'll create a blank dataset with each country and years 1989 through 2021
cdf <- map_df(countries, ~
                tibble(
                  country = (.),
                  year = rep(1989:2021)
                )) %>%
  full_join(ged2, by = c("country", "year")) %>%
  rename(osv = n) %>%
  mutate(osv = ifelse(is.na(osv), 0, osv)) %>% # Using the BASE ifelse rather than dplyr if_else
  full_join(ged3, by = "country") %>%
  rename(COWcode = country_id)
```

Now, let's assume that one-sided violence might be a function of whether an election is occurring and the GDP.

Bring in the V-Dem data
```{r Import the V-Dem data}
# Import the V-Dem data
vdem <- read_csv(file = here::here("Data sets", "vdem", "vdem" ,"V-Dem-CY-Full+Others-v12.csv")) %>%
  dplyr::select(COWcode, year, v2pepwrgen, e_gdp) %>%
  glimpse()

# Put these data into the cdf dataset
cdf <- cdf %>%
  full_join(vdem, by = c("COWcode", "year")) %>%
  rename(gender = v2pepwrgen,
         gdp = e_gdp) %>%
  mutate(lgdp = log(gdp + 1)) %>%
  glimpse()
```

Alright, let's try out regressing osv on gender and GDP
```{r Regressions}
# Running a core Poisson model
p4 <- glm(osv ~ gender + lgdp,
          family = poisson,
          data = cdf)

# Running a Zero-Inflated Poisson model using the PLSC library
p5 <- pscl::zeroinfl(osv ~ gender + lgdp,
                     dist = c("poisson"),
                     data = cdf)

# Moving to NEGATIVE BINOMIAL with the MASS library
p6 <- glm.nb(osv ~ gender + lgdp,
             data = cdf)


# Let's compare these models
modelsummary(models = list(p4, p5, p6),
             stars = TRUE)
```


Which of these models best predicts the outcome?


So let's build on THAT model

```{r Predicted counts}

# Let's plot out the marginal effects
marginaleffects::predictions(model = p6,
                             newdata = datagrid(gender = seq(-1, 1, by = 0.1))) %>%
  ggplot(aes(x = gender)) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high),
              fill = "snow") +
  geom_line(aes(y = predicted),
            size = 0.75) +
  geom_line(aes(y = conf.low),
            linetype = "dashed",
            size = 0.5) +
  geom_line(aes(y = conf.high),
            linetype = "dashed",
            size = 0.5) +
  theme(panel.border = element_rect(fill = NA,
                                    color = "black",
                                    size = 1),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12)) +
  labs(title = "Predicted count of one-sided violence",
       x = "V-Dem measurement of gender power balance",
       y = "Predicted count of one-sided violence")





# What if we broke out predictions by overall GDP?
marginaleffects::predictions(model = p6,
                             newdata = datagrid(gender = seq(-1, 1, by = 0.1),
                                                lgdp = c((mean(cdf$lgdp, na.rm = TRUE) - sd(cdf$lgdp, na.rm = TRUE)),
                                                          mean(cdf$lgdp, na.rm = TRUE),
                                                          (mean(cdf$lgdp, na.rm = TRUE) + sd(cdf$lgdp, na.rm = TRUE))))) %>%
  ggplot(aes(x = gender,
             fill = factor(lgdp))) +
  geom_ribbon(aes(ymin = conf.low,
                  ymax = conf.high)) +
  geom_line(aes(y = predicted),
            size = 0.75) +
  geom_line(aes(y = conf.low),
            linetype = "dashed",
            size = 0.5) +
  geom_line(aes(y = conf.high),
            linetype = "dashed",
            size = 0.5) +
  theme(panel.border = element_rect(fill = NA,
                                    color = "black",
                                    size = 1),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 12),
        legend.position = "bottom") +
  labs(title = "Predicted count of one-sided violence",
       x = "V-Dem measurement of gender power balance",
       y = "Predicted count of one-sided violence")


```





